Globally-centered autocovariances in MCMC

### 2020-09-03 Paper/Submissions/JCGS/First/main.tex
Pushing submission to JCGS changes

\documentclass[11pt]{article}

@@

% This is a package to accept utf8 input.  I normally do not use it in my documents, but it was here by default in Overleaf.

\usepackage{amssymb,amsthm,amsmath}
\usepackage{commath,hyperref}
\usepackage{enumerate,xcolor}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{caption,setspace}
\usepackage{subcaption}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\dd}[1]{\mathrm{d}#1}
\newcommand{\ds}{\displaystyle}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\text{Var}}
\newcommand{\Cov}{\text{Cov}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\B}{\mathcal{B}}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

@@

\maketitle

\onehalfspacing
\begin{abstract}
Autocovariances are a fundamental quantity of interest in Markov chain Monte Carlo (MCMC) simulations with autocorrelation function (ACF) plots being an integral visualization tool for performance assessment. Unfortunately, for slow mixing Markov chains, the empirical autocovariance can highly underestimate the truth. For multiple-chain MCMC sampling,  we propose a globally-centered estimator of the autocovariance function (G-ACvF) that exhibits significant theoretical and empirical improvements. We show that the bias of the G-ACvF estimator is smaller than the bias of the current state-of-the-art.  The impact of this improved estimator is evident in three critical output analysis applications: (1) ACF plots, (2) estimates of the Monte Carlo asymptotic covariance matrix, and (3) estimates of the effective sample size. Under weak conditions, we establish strong consistency of an improved asymptotic covariance estimator, and obtain its large-sample bias and variance.  The performance of the new estimators is demonstrated through various examples.
\end{abstract}

@@

\begin{figure}[htbp]
\centering
   \includegraphics[width=.5\linewidth]{plots/var-acf_n1000.pdf}
     \includegraphics[width=.5\linewidth]{plots/var-acf_n10000.pdf} 
\caption{VAR. ACF (left) and G-ACF (right) for the second chain for $m = 5$. (Top) $n = 10^3$  and (bottom) $n =  10^4$. The red line is the true ACF.}
\label{fig:var-acf}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
      \includegraphics[width = 2.0in]{plots/var-frob.pdf} 
      \includegraphics[width = 2.0in]{plots/var-ess.pdf}         
    \caption{VAR. (Left) Running plot for logarithm of Frobenius norm of A-SV and G-SV estimator. (Right) Running plot for logarithm of $\widehat{\textrm{ESS}}/mn$ using A-SV and G-SV estimator.}
    \label{fig:var-frob_n_ess}
\end{figure}

@@

@@

% We believe that the students from different races experience different socio-political conditions which causes them to engage in a "selective networking". This might cause formation of clusters in the social network. 
We sample two parallel Markov chains starting from two students belonging to different races.
 % and study its impacts on the average features of their immediate social group, especially their grades. Due to cluster formation in the network, a Markov chain will land on students within a cluster more often than on students outside it. As a consequence, the sample averages for the two chains will differ for a relatively shorter chain length. 
% 
Figure~\ref{fig:magnolia-acf} shows the ACF plots for the \textit{grade} feature at two different simulation sizes for $m=2$ parallel Markov chains. For a shorter chain length, both the chains have explored different clusters and as a consequence, the local and global mean do not agree. Regardless, G-ACF displays a clear advantage over locally-centered ACF for a chain length of $n=100$ samples whereas the latter takes $n=1000$ samples to reach the truth.

% It is hypothesized that students from certain communities tend to engage more with students of same or other marginally represented communities. We believe that this "selective networking" might cause formation of clusters in the network wherein students within a cluster engage more with each other than with the students outside it. Given that the graph is well-connected, the Markov chains will eventually explore all the nodes.

\singlespacing
\bibliographystyle{apalike}
\bibliography{sample}
\end{document}

>>>>

\documentclass[12pt]{article}

@@

% This is a package to accept utf8 input.  I normally do not use it in my documents, but it was here by default in Overleaf.
\usepackage{amsmath}
\usepackage{graphicx}
%\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\usepackage{amssymb,amsthm}
\usepackage{commath,hyperref}
\usepackage{enumerate,xcolor}
\usepackage{natbib}
\usepackage{graphicx}
% \usepackage{caption,setspace}
\usepackage{subcaption}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\dd}[1]{\mathrm{d}#1}
\newcommand{\ds}{\displaystyle}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\text{Var}}
\newcommand{\Cov}{\text{Cov}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\B}{\mathcal{B}}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

@@

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Globally-centered autocovariances in MCMC}
  \author{Medha Agarwal\hspace{.2cm}\\
    Department of Mathematics and Statistics, IIT Kanpur\\
    and \\
    Dootika Vats \\
    Department of Mathematics and Statistics, IIT Kanpur}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Title}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Autocovariances are a fundamental quantity of interest in Markov chain Monte Carlo (MCMC) simulations with autocorrelation function (ACF) plots being an integral visualization tool for performance assessment. Unfortunately, for slow mixing Markov chains, the empirical autocovariance can highly underestimate the truth. For multiple-chain MCMC sampling,  we propose a globally-centered estimator of the autocovariance function (G-ACvF) that exhibits significant theoretical and empirical improvements. We show that the bias of the G-ACvF estimator is smaller than the bias of the current state-of-the-art.  The impact of this improved estimator is evident in three critical output analysis applications: (1) ACF plots, (2) estimates of the Monte Carlo asymptotic covariance matrix, and (3) estimates of the effective sample size. Under weak conditions, we establish strong consistency of an improved asymptotic covariance estimator, and obtain its large-sample bias and variance.  The performance of the new estimators is demonstrated through various examples.
\end{abstract}

\noindent%
{\it Keywords:}  Markov chain Monte Carlo, autocorrelation plots, spectral variance estimators, visualization
\vfill

\newpage
\spacingset{1.5} % DON'T change the spacing!

@@

\begin{figure}[htbp]
\centering
   \includegraphics[width=.60\linewidth]{plots/var-acf_n1000.pdf}
     \includegraphics[width=.60\linewidth]{plots/var-acf_n10000.pdf} 
\caption{VAR. ACF (left) and G-ACF (right) for the second chain for $m = 5$. (Top) $n = 10^3$  and (bottom) $n =  10^4$. The red line is the true ACF.}
\label{fig:var-acf}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
      \includegraphics[width = 2.3in]{plots/var-frob.pdf} 
      \includegraphics[width = 2.3in]{plots/var-ess.pdf}         
    \caption{VAR. (Left) Running plot for logarithm of Frobenius norm of A-SV and G-SV estimator. (Right) Running plot for logarithm of $\widehat{\textrm{ESS}}/mn$ using A-SV and G-SV estimator.}
    \label{fig:var-frob_n_ess}
\end{figure}

@@

\bigskip
\begin{center}
{\large\bf SUPPLEMENTARY MATERIAL}
\end{center}

\begin{description}

\item[R Code:] Contains R code that reproduces the simulations, figured, and tables (.zip file).

\end{description}

@@

% We believe that the students from different races experience different socio-political conditions which causes them to engage in a "selective networking". This might cause formation of clusters in the social network. 
We sample two parallel Markov chains starting from two students belonging to different races.

% 
Figure~\ref{fig:magnolia-acf} shows the ACF plots for the \textit{grade} feature at two different simulation sizes for $m=2$ parallel Markov chains. For a shorter chain length, both the chains have explored different clusters and as a consequence, the local and global mean do not agree. Regardless, G-ACF displays a clear advantage over locally-centered ACF for a chain length of $n=100$ samples whereas the latter takes $n=1000$ samples to reach the truth.

\bibliographystyle{apalike}
\bibliography{sample}
\end{document}
